---
title: "K-2 T2 Behavioral Data (RA Version: Exploration in R)"
author: "Cha Cha Pillai, Mira Guleri, Angie Wang"
date: "8/13/2020"
output: pdf_document
---

Text in these white spaces will appear in your rendered file as text.
<!--But this text won't appear!-->

# This creates a heading
## This is a smaller heading
### An even smaller heading!

Click the Knit button to render the file and see what the text looks like.        

```{r set up, include=FALSE}
# This is our first code chunk. In the curly braces / chunk header above, 
# 'r' (which is automatically inserted) tells us that this is an R code chunk.
# We then named this chunk 'set up,' and specified that the contents of the chunk 
# shouldn't appear in the rendered file with the 'include = FALSE' argument.

# First, we load the packages that we'll be using. The packages have to 
# be installed before you load them, which you can do using install.packages().

library(Hmisc)      # for correlation matrix
library(psych)      # for descriptive statistics
library(pastecs)    # for descriptive statistics
library(knitr)      # for knitting to pdf
library(kableExtra) # for making pretty tables
library(ggpubr)     # for arranging plots
library(corrplot)   # for making correlogram
library(forcats)    # for reordering factor levels
library(tidyverse)  # for data wrangling

# You'll also want to check what your working directory is, so R knows where 
# to look for files. Try getwd() and setwd(). Alternatively, you can go to the
# top menu bar on your computer > Session > Set Working Directory > Choose Directory.

```

# Background and Participants

K-2 T2

* Second time point collected remotely via Zoom, May-June 2020 (end of the school year)
* N = 53 (43 participants returning from T1)
* Grades: K, 1, 2
* Ages: 5 to 8

# Data Cleaning

Data file: K2_T2_BehavioralData_20200728.csv

* Selected raw scores for CTOPP, TOWRE, W-J, plus scaled scores for TOWRE.
* Did not remove any subjects.

```{r clean data, include=FALSE}
# Next, we read in the .csv file. We included additional arguments in read_csv():
# (1) Set col_names as TRUE so the first row of the input will be used as the 
# column names, and will not be included in the data frame. (2) Specify that the 
# string "N/A" in the data should be interpreted as a missing value.

# read in T2 data
b2_data <- read_csv("K2_T2_BehavioralData_20200728.csv", 
                    col_names = TRUE, 
                    na = "N/A")

# look at the class of every column in the dataframe
#sapply(b2_data, class)

# you can do something like this to change the class of a specific column
#b2_data$k2_t2_b_grade_year <- as.factor(b2_data$k2_t2_b_grade_year)

# view your dataframe
#View(b2_data)

# After checking that the data looks normal, we use pipes (%>%) for a series 
# of actions on our original data frame, b2_data. The cleaned result is 
# assigned to a new variable, b2_clean.

# look at existing column names
#colnames(b2_data)

# clean the data
b2_clean <- b2_data %>% 
  # rename columns (new_name = old_name)
  rename(pid = participant_id, 
         age_year = k2_t2_b_age_years,
         age_months = k2_t2_b_age_months,
         age_days = k2_t2_b_age_days,
         grade = k2_t2_b_grade_year,
         ctopp_c_raw = k2_t2_b_color_raw,
         ctopp_l_raw = k2_t2_b_lttr_raw,
         elision_raw = k2_t2_b_elision_raw,
         swe_raw = k2_t2_b_swe_raw_fin,
         pde_raw = k2_t2_b_pde_raw_fin,
         swe_age = k2_t2_b_swe_agescal,
         pde_age = k2_t2_b_pde_agescal,
         swe_grade = k2_t2_b_swe_grdscal,
         pde_grade = k2_t2_b_pde_grdscal,
         wj_raw = k2_t2_b_wj_corr_nb) %>% 
  # select columns we want
  select (pid, age_year, age_months, age_days, grade,
          ctopp_c_raw, ctopp_l_raw, elision_raw,
          swe_raw, pde_raw, swe_age, pde_age,
          swe_grade, pde_grade, wj_raw) %>% 
  # create new column for age as a decimal value
  mutate(age = (age_year + (age_months/12) + (age_days/365))) %>% 
  # reorder grade variable to show in this order for plotting
  mutate(grade = fct_relevel(grade,"K", "1", "2")) %>% 
  # de-select columns that we don't need anymore
  select(-age_months, -age_days) %>% 
  # reorder columns
  select(pid, grade, age, age_year, everything())

```

\newpage <!--putting this on a new page in the rendered file-->
# Extra Practice: Manipulating dataframes with dplyr verbs 
Here's a useful dplyr cheatsheet: https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf

### Exercise 1
Create a new data frame with just the participant id, age, grade, and CTOPP Elision data. Try reordering the columns, like putting grade ahead of everything else.

```{r dplyr extra practice #1}

# 1. Solution

b2_elision <- b2_clean %>%
  select(grade, pid, age, elision_raw)

b2_elision

```

### Exercise 2a
Create a new data frame with just 2nd grade W-J data. (Hint: Use `filter()`). You should end up with 19 subjects.

### Exercise 2b
Still working with the same data frame in 2a (Hint: Use piping!), pick out subjects who scored 60 or above. You should end up with 8 subjects.

```{r dplyr extra practice #2}
# 2a. Solution



# 2b. Solution



```

### Exercise 3
Create a new dataframe with 1st and 2nd grade data. (Hint: Use logical operators to get both grade levels)

```{r dplyr extra practice #3}
# 3. Solution


```

### Exercise 4a
The `arrange()` function orders rows by values of a column or columns (low to high). Create a new dataframe where you arrange the `b2_clean` data by age (low to high)

### Exercise 4b
Use `drop_na()` to drop rows (subjects) that have any NA values in their scores.

```{r dplyr extra practice #4}
# 4a. Solution


# 4b. Solution


```

\newpage 
# Descriptive Statistics 
```{r descriptive stats, echo=FALSE}

# descriptive stats
b2_clean_descr <- b2_clean %>%
  select(- pid, - grade, - age_year) %>% # remove unnecessary columns
  stat.desc() # using stat.desc() from pastecs package

# comment this section out if it's currently throwing an error (this is just
# formatting stuff for now):
# put the desc stats in a nice table in our rendered file
kable(b2_clean_descr, 
      digits = 3, # round to 3 digits
      format = "latex", 
      caption = "K-2 T2 Descriptive Statistics",
      booktabs = TRUE) %>%
  # more formatting, from the kableExtra package
  kable_styling(latex_options = c("striped", "scale_down", "hold_position")) 


# Click the "Knit" button to render the file and view the table of
# descriptive statistics


# Calculate means by grade, use for plots later
b2_clean_means <- b2_clean %>%
  select(-pid, -age_year) %>% 
  group_by(grade) %>% 
  summarise_all("mean", na.rm=TRUE)

# TO-DO: Put means by grade in table for rendered file

b2_clean_means_K <- b2_clean_means %>% filter(grade == "K")
b2_clean_means_1 <- b2_clean_means %>% filter(grade == "1")
b2_clean_means_2 <- b2_clean_means %>% filter(grade == "2")

```

\newpage
# Visualizations

We are using the ggplot2 package (part of the tidyverse) to visualize the data.

* Learn more about the package here: https://ggplot2.tidyverse.org/.

* Cheat Sheet: https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf

In most cases you will:

1. Start with `ggplot()`
2. Supply a dataset in the first argument of `ggplot()`
3. Supply aesthetic mapping with `aes()`
4. Add on a `geom` function (like `geom_point()` or `geom_histogram()`)
5. Add on other layers and specifications (like editing the names of your axes, color, etc.)

## Examples using CTOPP Elision data

The Elision raw score is the number of items answered correctly before reaching the ceiling rule (3 in a row incorrect).

Basic scatterplot of elision raw scores:

```{r ggplot elision example basic, echo=FALSE}

# 1. We start with ggplot(). 

# 2. We specify the dataset to use in the graph in the first argument
# of ggplot(), in this case, b2_elision.

# 3. The x and y arguments of aes() specify which variables to map to the 
# x and y axes. Here, we specify that we want age on the x-axis and elision 
# raw scores on the y-axis.

# 4. We add on the function geom_point() to add a layer of points to a plot

ggplot(b2_clean, aes(x = age, y = elision_raw)) + 
  geom_point() # scatterplot

# We get a basic scatterplot!

```

\newpage
To enhance our scatterplot, we add on more layers and specifications:

```{r ggplot elision example enhanced, echo=FALSE}

# Adding color = grade as an argument in aes() to color-code the points by grade-level
p_el_1 <- ggplot(b2_clean, aes(x = age, y = elision_raw, color = grade)) + 
  geom_point() +
  scale_x_continuous(breaks = seq(5, 9, by=0.5)) + # adjusting x axis
  labs(title = "CTOPP Elision Raw Scores", # adding titles and labels with labs()
       x = "Age",
       y = "Raw Score (# correct)",
       caption = "This is my caption.") +
  theme_classic() # changing overall theme

p_el_1

```


```{r ggplot elision other plots, echo=FALSE, message=FALSE}

# Make a CTOPP Elision Density Plot - plot densities of elision raw data by grade

p_el_2 <- ggplot(b2_clean, aes(x = elision_raw, fill = grade)) +
  #geom_histogram(aes(y = ..density..), position = "identity", alpha = 0.3) +
  #geom_histogram(aes(y = ..density..), color = "black", fill = "white") +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = elision_raw, color = grade),
             linetype = "dashed") +
  labs(title = "CTOPP Elision Raw Scores",
       x = "Raw Score (# correct)",
       y = "Density",
       caption = "Dotted lines indicate raw score means by grade") +
  theme_classic()

p_el_2


# Make a CTOPP Elision Histogram (of all scores)

p_el_3 <- ggplot(b2_clean, aes(x = elision_raw)) +
  geom_histogram(fill = "lightblue") +
  labs(title = "CTOPP Elision Raw Scores",
       x = "Raw Score (# correct)",
       y = "Count") +
  theme_classic()

p_el_3


# Make a CTOPP Elision Box Plot - by grade

p_el_4 <- ggplot(b2_clean, aes(grade, elision_raw), fill = grade) +
  geom_boxplot(aes(fill = grade), alpha = 0.3) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(title = "CTOPP Elision Raw Scores",
       x = "Grade",
       y = "Raw Score (# Correct)") +
  theme_classic()

p_el_4

```

\newpage
## CTOPP Rapid Naming

### Rapid Color Naming

The Color Naming raw score is the time (in seconds) it takes to name all the colors on the page.

```{r ctopp color naming plots, echo=FALSE}

# Cha Cha

# TO-DO: Scatter (x = Age and y = Color Raw)


# TO-DO: Density (Color Raw Scores)


```

### Rapid Letter Naming

The Letter Naming raw score is the time (in seconds) it takes to name all the letters on the page.

```{r ctopp letter naming plots, echo=FALSE}

# Mira

# TO-DO: Scatter (x = Age and y = Letter Raw)
agelet_scat <- ggplot(b2_clean, aes(x = age, y = ctopp_l_raw)) + 
  geom_point(aes(color = age)) +
  geom_smooth(method = "lm", se = T, linetype = "dashed", color = "grey") + 
  labs(title = "Age and Letter Score (Raw)", 
       x = "Age (in years)",
       y = "Letter Raw (# read correctly)",
       caption = paste("r = ", 
                       round(cor(b2_clean$swe_raw, b2_clean$pde_raw, use="complete.obs"), 2), 
                       "(excludes 3 subjects with NAs)")) +
  theme_classic()
  agelet_scat

# TO-DO: Density (Letter Raw Scores)
let_den <- ggplot(b2_clean, aes(x = ctopp_l_raw, fill = grade)) +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = ctopp_l_raw, color = grade),
             linetype = "dashed") +
  labs(title = "CTOPP-2 (Rapid Letter Naming) Raw Scores",
       x = "Raw Score (# correct)",
       y = "Density",
       caption = "Dotted lines indicate raw score means by grade") +
  theme_classic()
let_den

```

### CTOPP Comparisons

```{r ctopp comparisons, echo=FALSE, message=FALSE, warning=FALSE}

# TO-DO: Scatter (Color and Letter)


# TO-DO: Scatter (Letter and Elision)


# TO-DO: Scatter (Color and Elision)

  
```

\newpage
## TOWRE

### Sight Word Efficiency (SWE)

The SWE raw score is the number of words read correctly in 45 seconds.

```{r towre-swe plots, echo=FALSE}

# Cha Cha

# TO-DO: Scatter (x = Age and y = Raw Scores)


# TO-DO: Density (Raw Scores)


```

### Phonemic Decoding Efficiency (PDE)

The PDE raw score is the number of words read correctly in 45 seconds.

```{r towre-pde plots, echo=FALSE}

# Mira

# TO-DO: Scatter (x = Age and y = Raw Scores)

agepde_scat <- ggplot(b2_clean, aes(x = age, y = pde_raw)) + 
  geom_point(aes(color = age)) +
  geom_smooth(method = "lm", se = T, linetype = "dashed", color = "grey") + 
  labs(title = "Age and Raw Letter Score", 
       x = "Age (in years)",
       y = "Letter Raw Score (letters read correctly)",
       caption = paste("r = ", 
                       round(cor(b2_clean$swe_raw, b2_clean$pde_raw, use="complete.obs"), 2), 
                       "(excludes 3 subjects with NAs)")) +
  theme_classic()

# TO-DO: Density (Raw Scores)


```

### TOWRE Comparisons
```{r towre comparisons, echo=FALSE, message=FALSE, warning=FALSE}

# Scatter (SWE and PDE Raw)

p_swe_pde <- ggplot(b2_clean, aes(x = swe_raw, y = pde_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "dashed", color = "grey") + 
  labs(title = "TOWRE SWE and PDE (raw scores)", 
       x = "SWE Raw Score (words read correctly)",
       y = "PDE Raw Score (nonwords read correctly)",
       caption = paste("r = ", 
                       round(cor(b2_clean$swe_raw, b2_clean$pde_raw, use="complete.obs"), 2), 
                       "(excludes 3 subjects with NAs)")) +
  theme_classic()

p_swe_pde

p_swe_pde2 <- ggplot(b2_clean, aes(x = swe_raw, y = pde_raw, color = grade)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, linetype = "solid") + 
  labs(title = "TOWRE SWE and PDE (raw scores)", 
       x = "SWE Raw Score (words read correctly)",
       y = "PDE Raw Score (nonwords read correctly)",
       caption = paste("r = ", 
                       round(cor(b2_clean$swe_raw, b2_clean$pde_raw, use="complete.obs"), 2), 
                       "(excludes 3 subjects with NAs)")) +
  theme_classic()

p_swe_pde2

```

\newpage
## W-J Letter-Word ID

The W-J raw score is the number of items answered correctly before reaching the ceiling rule (6 highest items that were administered are incorrect)

```{r w-j plots, echo=FALSE}

# TO-DO: Scatter (x = Age, y = Raw Score)


# TO-DO: Density


```

\newpage
## Correlations

Pasting initial T2 corrplot stuff here for now

TO-DO: T1-T2 correlations

```{r correlations, echo=FALSE}

p_all_corr <- b2_clean %>% 
  select(-c(pid, grade, age_year, swe_age, swe_grade, pde_age, pde_grade)) %>% 
  mutate(ctopp_c_raw = ctopp_c_raw * -1,
         ctopp_l_raw = ctopp_l_raw * -1) %>% 
  cor(use = "complete.obs") %>% 
  corrplot(method = "color")

```

