---
title: "K2_b_analysis_RA"
author: "Cha Cha Pillai, Mira Guleri, Angie Wang"
date: "7/31/2020"
output: pdf_document
---

Text in these white spaces will appear in your markdown doc as text.
<!--But this text won't appear!-->

# This creates a heading
## This is a smaller heading
###### An even smaller heading

Click the Knit button to render the file and see what the text looks like.        

```{r set up, include=FALSE}
# This is our first code chunk. In the curly braces / chunk header above, 
# 'r' (which is automatically inserted) tells us that this is an R code chunk.
# We then named this chunk 'set up,' and specified that the contents of the chunk 
# shouldn't appear in the rendered file with the 'include = FALSE' argument.

# First, we load the packages that we'll be using. The packages have to 
# be installed before you load them, which you can do using install.packages().

library(Hmisc)      # for correlation matrix
library(psych)      # for descriptive statistics
library(pastecs)    # for descriptive statistics
library(knitr)      # for knitting to pdf
library(kableExtra) # for making pretty tables
library(ggpubr)     # for arranging plots
library(corrplot)   # for making correlogram
library(forcats)    # for reordering factor levels
library(tidyverse)  # for data wrangling

# You'll also want to check what your working directory is, so R knows where 
# to look for files. Try getwd() and setwd(). Alternatively, you can go to the
# top menu bar on your computer > Session > Set Working Directory > Choose Directory.

```


```{r clean data, include=FALSE}
# Next, we read in the .csv file. We included additional arguments in read_csv():
# (1) Set col_names as TRUE so the first row of the input will be used as the 
# column names, and will not be included in the data frame. (2) Specify that the 
# string "N/A" in the data should be interpreted as a missing value.

# read in T2 data
b2_data <- read_csv("K2_T2_BehavioralData_20200728.csv", 
                    col_names = TRUE, 
                    na = "N/A")

# look at the class of every column in the dataframe
#sapply(b2_data, class)

# you can do something like this to change the class of a specific column
#b2_data$k2_t2_b_grade_year <- as.factor(b2_data$k2_t2_b_grade_year)

# view your dataframe
#View(b2_data)

# After checking that the data looks normal, we use pipes (%>%) for a series 
# of actions on our original data frame, b2_data. The cleaned result is 
# assigned to a new variable, b2_clean.

# look at existing column names
#colnames(b2_data)

# clean the data
b2_clean <- b2_data %>% 
  # rename columns (new_name = old_name)
  rename(pid = participant_id, 
         age_year = k2_t2_b_age_years,
         age_months = k2_t2_b_age_months,
         age_days = k2_t2_b_age_days,
         grade = k2_t2_b_grade_year,
         ctopp_c_raw = k2_t2_b_color_raw,
         ctopp_l_raw = k2_t2_b_lttr_raw,
         elision_raw = k2_t2_b_elision_raw,
         swe_raw = k2_t2_b_swe_raw_fin,
         pde_raw = k2_t2_b_pde_raw_fin,
         swe_age = k2_t2_b_swe_agescal,
         pde_age = k2_t2_b_pde_agescal,
         swe_grade = k2_t2_b_swe_grdscal,
         pde_grade = k2_t2_b_pde_grdscal,
         wj_raw = k2_t2_b_wj_corr_nb) %>% 
  # select columns we want
  select (pid, age_year, age_months, age_days, grade,
          ctopp_c_raw, ctopp_l_raw, elision_raw,
          swe_raw, pde_raw, swe_age, pde_age,
          swe_grade, pde_grade, wj_raw) %>% 
  # create new column for age as a decimal value
  mutate(age = (age_year + (age_months/12) + (age_days/365))) %>% 
  # reorder grade variable to show in this order for plotting
  mutate(grade = fct_relevel(grade,"K", "1", "2")) %>% 
  # de-select columns that we don't need anymore
  select(-age_months, -age_days) %>% 
  # reorder columns
  select(pid, grade, age, age_year, everything())

```

\newpage <!--putting this on a new page in the rendered file-->
# Extra Practice: Manipulating dataframes with dplyr verbs 
Here's a useful dplyr cheatsheet: https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf

1\. Create a new data frame with just the participant id, age, grade, and CTOPP Elision data. Try reordering the columns, like putting grade ahead of everything else.

```{r dplyr extra practice #1}
# 1. Solution

b2_elision <- b2_clean %>%
  select(grade, pid, age, elision_raw)

```

2a\. Create a new data frame with just 2nd grade W-J data. (Hint: Use filter()). You should end up with 19 subjects.

2b\. Still working with the same data frame in 2a (Hint: Use piping!), pick out subjects who scored 60 or above. You should end up with 8 subjects.

```{r dplyr extra practice #2}
# 2a. Solution



# 2b. Solution



```

3\. Create a new dataframe with 1st and 2nd grade data. (Hint: Use logical operators to get both grade levels)

```{r dplyr extra practice #3}
# 3. Solution


```

4a\. The arrange() function orders rows by values of a column or columns (low to high). Create a new dataframe where you arrange the b2_clean data by age (low to high)

4b\. Use drop_na() to drop rows (subjects) that have any NA values in their scores.

```{r dplyr extra practice #4}
# 4a. Solution


# 4b. Solution


```

\newpage 
# Descriptive Statistics 
```{r descriptive stats, echo=FALSE}

# descriptive stats
b2_clean_descr <- b2_clean %>%
  select(- pid, - grade, - age_year) %>% # remove unnecessary columns
  stat.desc() # using stat.desc() from pastecs package

# put the desc stats in a nice table in our rendered file
kable(b2_clean_descr, 
      digits = 3, # round to 3 digits
      format = "latex", 
      caption = "K-2 T2 Descriptive Statistics",
      booktabs = TRUE) %>%
  # more formatting, from the kableExtra package
  kable_styling(latex_options = c("striped", "scale_down", "hold_position")) 


# Click the "Knit" button to render the file and view the table of
# descriptive statistics

```

\newpage
# Visualizations
```{r ctopp plots, echo=FALSE}

#TO-DO: Angie make structure for plots

```

